Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates

### 2025-04-28 stochastic_vehicle_dynamics.tex
[MM] position change for adherence constraint description

>>>>

\subsection{Robust adherence limit constraint formulation}
The model employs a tire formulation that neglects combined slip effects, i.e., it does not account for the simultaneous utilization of longitudinal and lateral tyre forces. As a result, the optimal control problem must include an additional constraint to ensure that the total ground reaction forces remain within the bounds of the tyre's adherence ellipse.

The base version of the constraint for each axle can be expressed as a function of $\bx$ and $\bu$, as follows:
\begin{equation}
	h^\textrm{A}_j(\bx,\bu) = \left[ \left( \frac{X_j(\bx,\bu)}{\mu_{x,j}} \right)^2+ \left( \frac{Y_j(\bx,\bu)}{\mu_{y,j}}\right)^2\right]\frac{1}{Z_j^2(\bx,\bu)}-1\leq0\qquad j=1,2.\label{eq:adherence_constraint}
\end{equation}
where the superscript A denotes adherence, and the subscript $j$ indicates the axle index, with $j=1$ referring to the front axle and $j=2$ to the rear axle. Here, $X_j$ and $Y_j$ denote the longitudinal and lateral components of the in-plane ground forces, respectively, as illustrated in Figure~\ref{fig:vehicle_model}, while $Z_j$ represents the total vertical load acting on the axle. 
As can be seen from~\eqref{eq:adherence_constraint}, the points $\left(X_j,Y_j\right)$ are constrained to lie within an ellipse whose semi-axes are given by $\mu_{x,j}Z_j$ and $\mu_{y,j}Z_j$.

The gradient of the constraint w.r.t. the state vector $\bx$, denoted by $\na_{\bx}h^\textrm{A}_j(\bmu)$, is used to evaluate the standard deviation $\sigma^\textrm{A}_j$ associated with the constraint. This gradient can be computed using the chain rule, as follows:
\begin{equation}
	\na_{\bx}h^\textrm{A}_j(\bmu,\bu) = 
	\frac{\pd h^\textrm{A}_j}{\pd X_j}\at_{\bmu,\bu}\na_{\bx}X_j(\bmu,\bu) + 
	\frac{\pd h^\textrm{A}_j}{\pd Y_j}\at_{\bmu,\bu}\na_{\bx}Y_j(\bmu,\bu) + 
	\frac{\pd h^\textrm{A}_j}{\pd Z_j}\at_{\bmu,\bu}\na_{\bx}Z_j(\bmu,\bu)
\end{equation}
where the first element is zero because $X_j$ is a function of the input vector $\bu$ only.

For example, introducing a back-off term $\be=0.1$ in the constraint implies that the axle's adherence capacity can be exploited up to 90\% of its limit. This allows the optimizer to determine the most appropriate trade-off between the longitudinal force $X$ and the lateral force $Y$ --- often by reducing both components to some extent --- based on the specific requirements of the manoeuvre.
